---
title: "Antibiotics and mortality"
format: 
  html:
    self-contained: true
    toc: true
---

## Load packages

```{r, echo = FALSE}
# Load packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(data.table, ggplot2, haven, fixest, collapse, bacondecomp, did, ggrepel)

## Clear memory
rm(list=ls())

# Set seed 
set.seed(1234)

# Disable scientific notation
options(scipen=999)
```

# Examine real mortality data from AEJ-Applied paper

## View time-series

```{r, load-data}
# State-level data for regressions
state_mortality = as.data.table(haven::read_dta("data/raw/113743-V1/AEJApp-20080229_Datasets/20080229_state_data.dta"))

    # Make sure things we want to look at are in the same units
    state_mortality[, infl_pneumonia_rate := infl_pneumonia_rate*100000]
    state_mortality[, tb_rate := tb_rate*100000]

# National level data for image 
national_mortality =  as.data.table(haven::read_dta("data/raw/113743-V1/AEJApp-20080229_Datasets/20080229_national_data.dta"))

```

Quick look at the data

```{r, time-series-plot}
flu_plot = ggplot(data = subset(national_mortality, year >= 1928), 
                  aes(x = year, y = (influenza_pneumonia_total))) + 
    geom_point(size = 3, alpha = .75) + 
    theme_classic(base_size = 18) +
    geom_vline(xintercept = 1937, linetype="dashed") +
    labs(title = "Influenza and pneumonia mortality rate across time", 
         y = "", 
         subtitle = "Death rate per 100,000",
         x = "Year") +
    theme(legend.position = "none") 

flu_plot
```

```{r, save-time-series-plot, echo = FALSE}
ggsave(plot = flu_plot, 
       filename = "output/flu-mortality-1928-1950.pdf", 
      height = 6,
      width = 9,
      units = "in")
```


## Simple DiD

Compare changes in influenza death rates, before and after the invention of sulfa-drugs, to changes in tuberculosis death rates. Original AEJ paper discusses how TB was less affected. 

### Time-series comparison

```{r compare-time-series-to-tb, echo = FALSE}
national_mortality[,one:=1]
national_mortality[,two:=2]

ggplot(data = subset(national_mortality, year >= 1928), aes(x = year, y = (influenza_pneumonia_total), shape = as.factor(one))) + 
    geom_point(size = 3, alpha = .75) + 
    theme_classic(base_size = 14) +
    geom_vline(xintercept = 1937, linetype="dashed") +
    labs(title = "Influenza and pneumonia mortality rate across time", 
         y = "", 
         subtitle = "Death rate per 100,000",
         x = "Year") +
    theme(legend.position = "none") +
    geom_point(data = subset(national_mortality, year >= 1928), aes(x = year, y = (tuberculosis_total), shape = as.factor(two)), size = 3, alpha = .75) +
        theme_classic(base_size = 14) +
    labs(title = "Mortality rate across time", 
         y = "", 
         subtitle = "Death rate per 100,000",
         x = "Year") +
      scale_shape_manual(name = "", values=c(16, 17), labels = c("Influenza/pneumonia", "Tuberculosis"),  guide = 'legend') + 
  theme(legend.position = c(.8, 1))
```

### DiD comparison

To make this comparison, I will first work with the state-level data. We will reshape the data, call treatment 1937 and after, create post $\times$ treat variable. 

```{r, reshape-state-data, echo = FALSE}

# Reshape the data
state_mortality_reshape = melt(state_mortality, id.vars = c("state_pc", "year"),
                measure.vars = c("tb_rate", "infl_pneumonia_rate"))

# Create treatment variables
state_mortality_reshape[, treated_disease := ifelse(variable == "tb_rate", 0, 1)]
state_mortality_reshape[, treated_time := ifelse(year >= 1937, 1, 0)]
state_mortality_reshape[, treated := treated_time*treated_disease]

```

Run simple DiD

```{r, simple-did}
simple_did = fixest::feols(value  ~ treated | treated_disease + state_pc + year, 
              data = state_mortality_reshape)
etable(simple_did)

effect = simple_did$coeftable[1,1]
```


There is a `r round(effect, 1)` change in the influenza and pneumonia mortality rate per 100,000, relative to the change in tuberculosis mortality, following the advent of sulfa-drugs. 


# Example using fake data

Add random variable to each state to get fake treated_year

First create percentiles

```{r}
quantiles = state_mortality[year == 1928, .(untreated = quantile(infl_pneumonia_rate, 0.25, na.rm = TRUE), low = quantile(infl_pneumonia_rate, 0.5, na.rm = TRUE), high = quantile(infl_pneumonia_rate, 0.75, na.rm = TRUE))]
```

Use these percentiles to create cutoffs in fake treatment.

-   Lowest 25% is "never treated"

-   25-50% is late

-   50-75% is middle

-   75%+ is early

```{r}

state_mortality[year == 1928, fake_year_treated := ifelse(infl_pneumonia_rate <= quantiles$untreated, 2020, ifelse(infl_pneumonia_rate > quantiles$untreated & infl_pneumonia_rate <= quantiles$low, 1945,
                                     ifelse(infl_pneumonia_rate > quantiles$low & infl_pneumonia_rate <= quantiles$high, 1940, 1930)))]

state_mortality = state_mortality[, fake_year_treated := fake_year_treated[year == 1928], by = state_pc]

```

## Create data with staggered adoption, growing, and heterogenous treatment effect.

```{r}
# Keep only the variables that we need from 1928
timing_yes_growing_yes = state_mortality[year == 1928,.(state_pc, infl_pneumonia_rate, fake_year_treated, pop3)]

# Expand the dataset to be from 1928 until 1950
full_state_year_grid = data.table(expand.grid(year=1928:1950, state_pc =unique(timing_yes_growing_yes$state_pc)))

# Create error
error_mean = 0
error_sd = 10
error_scale = 1

full_state_year_grid[,rand_e:=1-rnorm(.N, 0, error_sd)]
full_state_year_grid[year == 1928, rand_st:=1-rnorm(.N, 0, error_sd)]
full_state_year_grid[state_pc == "AL", rand_yr:=1-rnorm(.N, 0, error_sd)]

full_state_year_grid = full_state_year_grid[, rand_st := rand_st[year == 1928], by = state_pc]
full_state_year_grid = full_state_year_grid[, rand_yr := rand_yr[state_pc == "AL"], by = year]

# Merge the two togethers
timing_yes_growing_yes = merge(timing_yes_growing_yes, full_state_year_grid, by = "state_pc")

# Add fake treatment status in each year
timing_yes_growing_yes[, treated := ifelse(year >= fake_year_treated, 1, 0)]

# Add "natural" slope from pre-trend
timing_yes_growing_yes[, fake_value := infl_pneumonia_rate - .1*rand_st + .05*rand_yr + .1*rand_e]

# Add treatment effect
timing_yes_growing_yes[, fake_value := ifelse(fake_year_treated == 1940, fake_value + 0*effect*treated + (.2*simple_did$coeftable[1,1]*(year - fake_year_treated)*treated), ifelse(fake_year_treated == 1945, fake_value + 0*effect*treated + (.1*simple_did$coeftable[1,1]*(year - fake_year_treated)*treated),fake_value + 0*effect*treated + (.3*simple_did$coeftable[1,1]*(year - fake_year_treated)*treated)))]

# Bound at zero
timing_yes_growing_yes[, fake_value := ifelse(fake_value < 0, 0, fake_value)]

# Calculate true treatment effect 
timing_yes_growing_yes[, true_treatment_effect := fake_value - (infl_pneumonia_rate - .1*rand_st + .05*rand_yr + .1*rand_e)]

# Collapse into timing groups
timing_yes_growing_yes_collapsed = collap(timing_yes_growing_yes, ~ year + fake_year_treated, fmean, cols = 10, w = timing_yes_growing_yes$pop3)


```
Determine true treatment effect, ATT by group, and avg-ATT by treated year
```{r}
# Now figure out 
    # the average treatment effect for each group
    # the average treatment effect on the treated. 

yearly_true_treatment_effect = timing_yes_growing_yes[, list(avg_true_treatment_effect = fmean(true_treatment_effect),
                                             count = fnobs(true_treatment_effect)),
                                      by = c("fake_year_treated", "treated", "year")]
yearly_true_treatment_effect[treated == 1, event_time := year - fake_year_treated]
annual_avg_treatment_effect = yearly_true_treatment_effect[, list(avg_true_treatment_effect = fmean(avg_true_treatment_effect, w = count), weight_for_total = fsum(count)),
                                      by = c("event_time")]


group_true_treatment_effect = timing_yes_growing_yes[, list(mean = fmean(true_treatment_effect),
                                             count = fnobs(true_treatment_effect)),
                                      by = c("fake_year_treated", "treated")]

group_att = group_true_treatment_effect[treated == 1]
```


Plot data

```{r}
timing_yes_growing_yes_plot = ggplot(data = timing_yes_growing_yes_collapsed,
                                    aes(x = year, y = (fake_value), shape = as.factor(fake_year_treated))) + 
    geom_point(size = 3, alpha = .66) + 
    theme_classic(base_size = 18) +
    geom_vline(xintercept = 1944.5, linetype="dashed") +
    geom_vline(xintercept = 1929.5, linetype="dashed") +
    geom_vline(xintercept = 1939.5, linetype="dashed") +
    labs(title = "", 
         y = "", 
         subtitle = "Influenza/pneumonia mortality death rate per 100,000",
         x = "Year") +
      scale_shape_manual(name = "", 
                         values=c(15, 16, 17, 4), 
                         labels = c("Early", 
                                    "Middle", 
                                    "Late", 
                                    "Never"),  
                         guide = 'legend') + 
  theme(legend.position = c(.9, .9))
timing_yes_growing_yes_plot
```

```{r, save-fake-time-series-plot, echo = FALSE}
ggsave(plot = timing_yes_growing_yes_plot, 
       filename = "output/fake-data-flu-mortality-1928-1950.pdf", 
      height = 6,
      width = 9,
      units = "in")
```


## Estimate ATT

### Two way fixed-effects
```{r}
timing_yes_growing_yes_twfe = fixest::feols(fake_value ~ treated | state_pc + year, data = timing_yes_growing_yes)

# Estimate of the treatment effect is 
timing_yes_growing_yes_twfe$coeftable[1,1]
timing_yes_growing_yes_twfe
```

### Sun & Abraham (2020)

```{r}
timing_yes_growing_yes_sa = feols(fake_value ~ sunab(fake_year_treated, year)  ## Our key interaction: time × treatment status
		  |                    ## Other controls
		  state_pc + year,                             ## FEs
		 cluster = ~state_pc,                          ## Clustered SEs
		 data = timing_yes_growing_yes)
sa_output = summary(timing_yes_growing_yes_sa, agg = "ATT")
sa_output
```
### Stacked

Not sure of R command for this. So I am exporting to stata. 
```{r}
fwrite(timing_yes_growing_yes, 
       file = "timing_yes_growing_yes.csv")
```

Read in ATT and Event study
```{r}
timing_yes_growing_yes_stacked_att = fread("timing_yes_growing_yes_stacked_att.csv")
timing_yes_growing_yes_stacked_es = fread("timing_yes_growing_yes_stacked_es.csv")
```

### Callaway Sant'Anna

```{r}
# Make a copy so we don't override. 
timing_yes_growing_yes_cs = copy(timing_yes_growing_yes)

# Create numeric state-group
timing_yes_growing_yes_cs[,state_group := as.numeric(as.factor(state_pc))]

# Never treated should be set to zero
timing_yes_growing_yes_cs[,fake_year_treated := ifelse(fake_year_treated == 2020, 0, fake_year_treated)]

# Run CS
timing_yes_growing_yes_attgt = att_gt(yname = "fake_value",
                        tname = "year",
                        idname = "state_group",
                        gname = "fake_year_treated",
                        xformla = ~1,
                        data = timing_yes_growing_yes_cs
                        )

timing_yes_growing_yes_cs = aggte(timing_yes_growing_yes_attgt, type = "dynamic")
summary(timing_yes_growing_yes_cs)
```

## Plot the three ATT comparisons next to one another. 

```{r}


 # Create storage dataset. 
  storage_temp = data.table(order = numeric(), b=numeric(), se=numeric(), Method=character())
  
  # Add TWFE results to storage dataset
    temp = as.data.table(
            cbind(
                1, 
                timing_yes_growing_yes_twfe$coeftable[1,1], 
                timing_yes_growing_yes_twfe$coeftable[1,2], 
                "TWFE"
                )
            )
    storage_temp = rbind(storage_temp, temp, 
                                           use.names = FALSE)
    
    # Add SA results to storage dataset
    temp = as.data.table(
            cbind(
                2, 
                sa_output$coeftable[,1], 
                sa_output$coeftable[,2], 
                "Sun & Abraham (2020)"
                )
            )
    storage_temp = rbind(storage_temp, temp, 
                                           use.names = FALSE)
    
      

    # Add CS results to storage dataset
    temp = as.data.table(
            cbind(
                3, 
                timing_yes_growing_yes_cs$overall.att, 
                timing_yes_growing_yes_cs$overall.se, 
                "Callaway Sant'Anna"
                )
            )
    storage_temp = rbind(storage_temp, temp, 
                                           use.names = FALSE)
    
    # Add Stacked results to storage dataset
    temp = as.data.table(
            cbind(
                4, 
                timing_yes_growing_yes_stacked_att$b[1], 
                timing_yes_growing_yes_stacked_att$se[1], 
                "Stacked"
                )
            )
    timing_yes_growing_yes_att_data = rbind(storage_temp, temp, 
                                           use.names = FALSE)

    # Ensure storage is numeric where it should be
    timing_yes_growing_yes_att_data[, b := as.numeric(b)]
    timing_yes_growing_yes_att_data[, se := as.numeric(se)]
    timing_yes_growing_yes_att_data[, order := as.numeric(order)]

```

Plot the ATTs next to one another. 
```{r, plot-att}
    # Add truth 
    true_avg_att = annual_avg_treatment_effect[event_time > 0, list(avg_true_treatment_effect = fmean(avg_true_treatment_effect, w = weight_for_total))]
    true_avg_att = mean(true_avg_att$avg_true_treatment_effect)
timing_yes_growing_yes_att_plot = ggplot(data = timing_yes_growing_yes_att_data, 
       aes(x = factor(order), 
           y = b, 
           ymin = b - 1.96*se, 
           ymax = b + 1.96*se,
           shape = Method, 
           color = Method)) + 
    geom_point(size=3) +
    geom_errorbar(width=0.05) +
    theme_minimal() +
    geom_hline(yintercept = 0) +
    geom_hline(yintercept = true_avg_att, linetype = "dashed") +
    annotate("text", label = "True ATT", x = 4.4, y = true_avg_att*.95, size = 4)   +
    labs(title = "ATT estimates by method", 
         y = "", 
         subtitle = "Change in influenza/pneumonia mortality death rate per 100,000", 
         x = "")  + 
    scale_x_discrete(breaks=c("1","2","3", "4"),
        labels=c("TWFE", "Sun & Abraham (2020)", "Callaway Sant'Anna", "Stacked")) +
    theme_classic(base_size = 18) +
theme(legend.position = "none")
timing_yes_growing_yes_att_plot
```
```{r, save-fake-att-comparison, echo = FALSE}
ggsave(plot = timing_yes_growing_yes_att_plot, 
       filename = "output/fake-data-att-comparison.pdf", 
      height = 6,
      width = 9,
      units = "in")
```

### Bacon-decomposition to confirm ATT from TWFE

```{r}
timing_yes_growing_yes_bacon = bacon(fake_value ~ treated,
                  data = as.data.frame(timing_yes_growing_yes),
                  id_var = "state_pc",
                  time_var = "year")
timing_yes_growing_yes_bacon = as.data.table(timing_yes_growing_yes_bacon)
timing_yes_growing_yes_bacon[, untreated := ifelse(untreated == 99999, "Never", untreated)]

timing_yes_growing_yes_bacon[, label := do.call(paste, c(.SD, sep = " vs. ")), .SDcols = c("treated", "untreated")]

# Show bacon decomp 
timing_yes_growing_yes_bacon

# Now compare this to the group-specific real ATT
group_att

# weighted average of the decomposition, this should be the same as TWFE DiD
sum(timing_yes_growing_yes_bacon$estimate * timing_yes_growing_yes_bacon$weight)
```
```{r}
att_1930 = unlist(group_att[fake_year_treated == 1930, .(mean)])
att_1940 = unlist(group_att[fake_year_treated == 1940, .(mean)])
att_1945 = unlist(group_att[fake_year_treated == 1945, .(mean)])
```
Plot the decomposition
```{r}
timing_yes_growing_yes_bacon_plot = ggplot(timing_yes_growing_yes_bacon) +
      geom_hline(yintercept = att_1930, linetype = "dashed") +
    geom_hline(yintercept = att_1940, linetype = "dashed") +
    geom_hline(yintercept = att_1945, linetype = "dashed") +
  aes(x = weight, y = estimate, shape = factor(type), label = label, color = factor(type)) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0) + 
  theme_minimal(base_size = 18) +
        geom_point(x = .2, y = att_1940, alpha = 0, size = 3) +
    geom_point(x = .2, y = -30, size = 3, alpha = 0) +
    geom_point(x = .2, y = -25, size = 3, alpha = 0) +
    annotate("text", label = "True 1930", x = .2, y = att_1930*.95, size = 4)   +
    annotate("text", label = "True 1940", x = .205, y = -32, size = 4)    +
    annotate("text", label = "True 1945", x = .205, y = -11, size = 4)    +
          geom_label_repel(force = 10, max.time = 10, show.legend = FALSE, seed = 1234) +
  labs(x = "Weight", y = "Estimate", shape = "Type", color = "Type", title = "Bacon decomposition") +
     guides(col = guide_legend(ncol = 3)) +
    theme(legend.position="bottom", 
          legend.title = element_blank())


 
  

timing_yes_growing_yes_bacon_plot
```


```{r, echo = FALSE}
ggsave(plot = timing_yes_growing_yes_bacon_plot, 
       filename = "output/bacon-decomposition.pdf", 
      height = 6,
      width = 9,
      units = "in")
```

## Event-study

```{r event-study-function, echo = FALSE, eval = TRUE}

es_plot = function(dt, id_var, time_var, year_treated_var, value_var, untreated_value, ref_value = -1) {
  # Create new temp data.table
  dt_temp = copy(dt)
  
  # Add event-time
  dt_temp[, event_time := get(time_var) - get(year_treated_var)]
  dt_temp[, event_time := ifelse(get(year_treated_var) == untreated_value, ref_value,  event_time)]

  # Add ever treated
  dt_temp[, ever_treated := ifelse(get(year_treated_var) != untreated_value, 1,  0)]
  
  # Create storage dataset. 
  storage_temp = data.table(event_time = numeric(), b=numeric(), se=numeric(), Method=character())
  
  # Run TWFE
  formula = as.formula(paste0(value_var, " ~ i(event_time, ever_treated, ref = ", ref_value, ")",
                               " | ", id_var, " + ", time_var))

  did_temp = fixest::feols(formula, data = dt_temp)
  
  # Add results to storage dataset
    temp = as.data.table(
            cbind(
                did_temp$model_matrix_info[[1]]$items, 
                append(did_temp$coefficients, 0, abs(min(did_temp$model_matrix_info[[1]]$items))-1), 
                append(did_temp$se, 0, abs(min(did_temp$model_matrix_info[[1]]$items))-1), 
                rep("TWFE",each=length(did_temp$model_matrix_info[[1]]$items))
                )
            )
    storage_temp = rbind(storage_temp, temp, 
                                           use.names = FALSE)
    
  # Run Sun and Abraham
  formula_sa = as.formula(paste0(value_var, " ~ sunab(" , year_treated_var, ", ", time_var, ")",
                               " | ", id_var, " + ", time_var))
  
  sa_temp = fixest::feols(formula_sa, data = dt_temp)
    
  # Extract coefs
  coef_sa_temp = sa_temp$coeftable
  
  # Add results to storage dataset
  temp = as.data.table(
            cbind(
                sa_temp$model_matrix_info[[1]]$items, 
                append(coef_sa_temp[,1], 0, abs(min(sa_temp$model_matrix_info[[1]]$items))-1), 
                append(coef_sa_temp[,2], 0, abs(min(sa_temp$model_matrix_info[[1]]$items))-1), 
                rep("Sun & Abraham (2020)",
                    each=length(sa_temp$model_matrix_info[[1]]$items))
                )
            )
    storage_temp = rbind(storage_temp, temp, 
                                       use.names = FALSE)

    # Callaway Sant'Anna
    
    # Turn id var into numeric
    dt_temp[,state_group := as.numeric(as.factor(get(id_var)))]
    
    # Make sure that untreated year is set to zero
    dt_temp[ ,year_treated_temp := ifelse(get(year_treated_var) == untreated_value, 0, get(year_treated_var))]
    
 
    # Run CS
    example_attgt = att_gt(yname = value_var,
                        tname = time_var,
                        idname = "state_group",
                        gname = "year_treated_temp",
                        xformla = ~1,
                        data = dt_temp
                        )
    # Aggregate
    es = aggte(example_attgt, type = "dynamic")

    # Add to temp storage
    temp = as.data.table(
            cbind(
               es$egt, 
                es$att.egt, 
                es$se.egt, 
                rep("Callaway Sant'Anna", each=length(es$egt))
                )
            )
    
    storage_temp = rbind(storage_temp, temp, 
                                       use.names = FALSE)
    
    # Add Stacked temp storage
    temp = as.data.table(
            cbind(
               timing_yes_growing_yes_stacked_es$type, 
                timing_yes_growing_yes_stacked_es$b, 
                timing_yes_growing_yes_stacked_es$se, 
                rep("Stacked", each=length(timing_yes_growing_yes_stacked_es$se))
                )
            )
    
    storage_temp = rbind(storage_temp, temp, 
                                       use.names = FALSE)
    
    # Add truth 
       truth_to_add =  annual_avg_treatment_effect[event_time > 0, se := 0]
    temp = as.data.table(
            cbind(
               truth_to_add$event_time, 
                truth_to_add$avg_true_treatment_effect, 
                truth_to_add$se, 
                rep("True avg effect", each=length(es$egt))
                )
            )
    
    storage_temp = rbind(storage_temp, temp, 
                                       use.names = FALSE)
    
    
    # Ensure storage is numeric where it should be
    storage_temp[, b := as.numeric(b)]
    storage_temp[, se := as.numeric(se)]
    storage_temp[, event_time := as.numeric(event_time)]
    
    # Return value
    return(storage_temp)
}
```

Run TWFE, Sun Abraham, and CS event-studies, storing coefs and se. 

```{r, run-es-code}
timing_yes_growing_yes_es_data = es_plot(
    dt = timing_yes_growing_yes, 
    id_var = "state_pc",
    time_var = "year",
    year_treated_var = "fake_year_treated",
    value_var = "fake_value",
    untreated_value = 2020,
    ref_value = -1
)
```

Plot the event-studies on top of one another. 
```{r, plot-es}
timing_yes_growing_yes_es_plot = ggplot(data = timing_yes_growing_yes_es_data, 
       aes(x = event_time, 
           y = b, 
           ymin = b - 1.96*se, 
           ymax = b + 1.96*se,
           shape = Method, 
           color = Method)) + 
    geom_point(size=2) +
    geom_errorbar(width=0.1) +
    theme_minimal() +
    geom_vline(xintercept = -.5) +
    geom_hline(yintercept = 0) +
    labs(title = "Event-study estimates by method", 
         y = "", 
         subtitle = "Change in influenza/pneumonia mortality death rate per 100,000", 
         x = "Years since sulfa available in state")  + 
  theme(legend.position = c(.875, .75))
timing_yes_growing_yes_es_plot
```

```{r, save-es-series-plot, echo = FALSE}
ggsave(plot = timing_yes_growing_yes_es_plot, 
       filename = "output/timing_yes_growing_yes_es_plot.pdf", 
      height = 6,
      width = 9,
      units = "in")
```

Zoom into -8 to +8, where the majority of treated years are, to see how TWFE is different that other two methods. 

```{r, plot-es-zoom}
timing_yes_growing_yes_es_zoom_plot = timing_yes_growing_yes_es_plot + 
    xlim(-8.5, 8) +
    ylim(-35, 5)  + 
  theme(legend.position = c(.875, .7))

timing_yes_growing_yes_es_zoom_plot
```


```{r}
ggsave(plot = timing_yes_growing_yes_es_zoom_plot, 
       filename = "output/timing_yes_growing_yes_es_zoom_plot.pdf", 
      height = 6,
      width = 9,
      units = "in")
```


